Mutation loads in rheumatoid arthritis data from SComatic

### 2024-01-31 scomatic/reports/driver_coverage.Rmd
add infercnv rmd to show all plots, changes to driver coverage rmd

---
title: "Mutation loads in rheumatoid arthritis data from SComatic"
author: "Alexandra Tidd"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    fig_width: 8
    keep_md: true
    code_folding: show
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 4
    theme: lumen
params:
    rerun: true
---

```{r setup, include = F, message = F, warning = F, class.source = 'fold-hide'}
# wd <- ifelse(Sys.info()['nodename'] == "mib119104s", '~/Volumes/', '') %>% paste0('/lustre/scratch125/casm/team268im/at31/RA_som_mut/scomatic/') ; setwd(wd)
# rmarkdown::render('reports/driver_coverage.Rmd', output_file = 'driver_coverage.html', output_dir = 'reports/', params = list(rerun = T))

@@

```{r plot_drivers, class.source = "fold-hide", warning = F, message = F}
drivers %>%
  dplyr::distinct(celltype, gene, assay, qglobal_cv) %>%
  dplyr::mutate(neglog10q = -log10(qglobal_cv)) %>%
  ggplot(aes(x = tidytext::reorder_within(gene, -neglog10q, celltype), 
             y = neglog10q, fill = celltype)) +
  geom_col(position = 'dodge') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = -90, hjust = 0)) +
  facet_grid(assay ~ celltype, scales = 'free_x', space = 'free_x') +
  tidytext::scale_x_reordered() +
  labs(x = 'gene', y = '-log10(p)')
```

@@

```{r get_driver_pos}
# load biomart
library(biomaRt)

# load ref
ensembl <- useEnsembl(biomart = 'ensembl',
                      dataset = 'hsapiens_gene_ensembl')

# get positions
drivers_pos <-
  getBM(attributes = c('hgnc_symbol', 'chromosome_name', 'start_position', 'end_position'),
        filters = 'hgnc_symbol', values = unique(drivers$gene),
        mart = ensembl) %>%
  # fix colnames
  dplyr::mutate(gene = hgnc_symbol, 
                chr = chromosome_name, 
                start = start_position,
                end = end_position) %>%
  # remove weird chromosomes
  dplyr::filter(chr %in% c(as.character(1:22, 'X', 'Y')))

>>>>

---
title: "Mutation loads in rheumatoid arthritis data from SComatic"
author: "Alexandra Tidd"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    fig_width: 8
    keep_md: false
    code_folding: show
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 4
    theme: lumen
params:
    rerun: true
---

```{r setup, include = F, message = F, warning = F, class.source = 'fold-hide'}
# wd <- ifelse(Sys.info()['nodename'] == "mib119104s", '~/Volumes/', '') |> paste0('/lustre/scratch125/casm/team268im/at31/RA_som_mut/scomatic/') ; setwd(wd)
# rmarkdown::render('reports/driver_coverage.Rmd', output_file = 'driver_coverage.html', output_dir = 'reports/', params = list(rerun = F))

@@

```{r plot_drivers, class.source = "fold-hide", warning = F, message = F, fig.width = 10}
drivers %>%
  dplyr::distinct(celltype, gene, assay, qglobal_cv) %>%
  dplyr::mutate(neglog10q = -log10(qglobal_cv)) %>%
  ggplot(aes(x = tidytext::reorder_within(gene, -neglog10q, celltype), 
             y = neglog10q, fill = celltype)) +
  geom_col(position = 'dodge') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = -90, hjust = 0)) +
  facet_grid(assay ~ celltype, scales = 'free_x', space = 'free_x') +
  tidytext::scale_x_reordered() +
  labs(x = 'gene', y = '-log10(p)')
```

@@

```{r get_drivers_pos}
drivers_pos <- xfun::cache_rds({

  # load ref
  library(biomaRt)
  ensembl <- useEnsembl(biomart = 'ensembl', dataset = 'hsapiens_gene_ensembl')

  # get positions
  drivers_pos <-
    getBM(attributes = c('hgnc_symbol', 'chromosome_name', 'start_position', 'end_position'),
          filters = 'hgnc_symbol', values = unique(drivers$gene),
          mart = ensembl) %>%
    tibble::as_tibble() %>% 
    # fix colnames
    dplyr::mutate(gene = hgnc_symbol, 
                  chr = chromosome_name, 
                  start = start_position,
                  end = end_position) %>%
    # remove weird chromosomes
    dplyr::filter(chr %in% c(as.character(1:22, 'X', 'Y')))

  # return
  drivers_pos

}, file = 'drivers_pos.rds')
